---
title: "Assignment 1 ST302"
author: "your name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F)


```





## Question 1. 
## Describe the different steps in the evolution of the plot in the video.


Picture 1:
First picture is a boxplot. The code looks something like this:
ggplot(DATA, aes(x=region, y=student_ratio))+geom_boxplot() + labs(title = "The Evolution of a ggplot")

Picture 2:
The boxplot is grouped from the region with the highest student ratio to the lowest.

Picture 3:
Clear Change in vector. From top to bottom, it now shows  lowest student ratio to the highest.

Picture 4:
The x-axis changes from showing grids in groups of 20, to groups of 25

Picture 5:
Color is added to distinguish each continent from each other visually.

Picture 6:
Violin plots of Student Ratios in all regions. Again, Drawn on the horizontal axis.

Picture 7:
We can now see each individual sample from each continent. However, now the data is harder to read.

Picture 8:
Each circle has become transparent and we can now see where each sample is better. Especially in Europe.

Picture 10.:
A boxplot is overlaid on the data. It has only the outline and isn't filled in with color. However, the fact the data is still isn a straight line makes it hard to read individual values.

Picture 11:
The dots are spread out from each other but still gives the the correct student to teacher ratio. Allows us to more clearly see the individual samples.

Picture 12:
Boxplot is removed.

Picture 13:
Mean value of student to teacher ratio is displayed for each Continent.

Picture 14:
An X-Intercept line is added at around ~24.

Picture 15:
A line is drawn from each mean value to the X-Intercept line.

Picture 16:
Facts about the plot are added. E.G: Worldwide Average: 23.5 students per teacher, the Continental average and Countries per continent is listed, and the Central African Republic has by far the most students per teacher.

Picture 17:
Arrows are drawn from the facts to the relevant informaiton on the ggplot.

Picture 18:
The size of the arrows are made smaller.

Picture 19:
The Student to Teacher Ratio grid is changed back to grids of 20 except for the first grid, which is 1-20.

Picture 20:
The samples are shown as a map of the world, seperate to the recorded data. They are distinguished by color to indicate which continent.


## Question 2



a.
This is a Spiral Plot that shows the amount of Covid-19 Cases in the United States within the years 2020-2022.
The 0 to 150k cases shows the first amount of recorded cases starting within March of 2020.
You cannot accurately see the exact values of how much cases there is in a day/month, but you can get a general idea of when there were the most amount of cases.
For example, you can see that the highest amount of recorded cases was at the start of 2022 and the amount of cases are getting higher.
It is a very good plot for propaganda purposes as a spiral plot is more unique than a histogram or line chart, but I wouldn't recommend it among Staticians as it is difficult to get precise numbers.

    
b. 
Advantages of the Spiral Plot:
More eye-catching & attention grabbing than the more used, precise plots like bar charts and scatter plots.
Since it is such a distinctive plot., it helps draw people into the conversation about the topic the plot is about.


Disadvantages of the Spiral Plot:
Nearly impossible to get precise values on a given day/month.
It does not give minimum, maximum and the average amount of cases in a given month.

## Question 3

```{r, echo=TRUE}
library(tidyverse)
water <- read_csv(here::here("hwk", "Data","BathingWaterQuality2013.csv"))
water <- mutate(water,Date= lubridate::dmy(Date))
water4 <- filter(water, Beach %in% c("Sutton", "Portmarnock" ,"Skerries" ,"Howth - Claremont"))


```


a.Use dplyr count to check how many readings there are for each Beach.
install.packages("here")
```{r}
water %>% count(Beach)
```

b. Calculate the mean E_Coli and mean Enterococci for the beaches.
```{r}
water %>% group_by(Beach) %>% summarise(meanEcoli = mean(E_Coli), meanEnter = mean(Enterococci))
```

c. Using the four beaches in water4, plot Enterococci versus date, color by Beach. Use geom_point and geom_line(). Use geom_hline(yintercept=100, color="magenta") to show the guideline value.
```{r}
ggplot(data=water4, mapping = aes(x=Enterococci, y= Date, color= Beach))+
geom_point() + geom_line() + geom_hline(yintercept=100, color="magenta")

```

d. Repeat c. using E_Coli.
```{r}
ggplot(data=water4, mapping = aes(x=E_Coli, y= Date, color= Beach))+
geom_point() + geom_line() + geom_hline(yintercept=100, color="magenta")

```

e. Which beach has the highest readings?  If so which beach and when?
   Do any of the beaches exceed the guidelines? If so which beach and when? E_Coli = 2000(95%), 100(90%). Enterococci = 100(80%)
```{r}
water %>% filter(Enterococci == max(Enterococci))
water %>% filter(E_Coli == max(E_Coli))

water %>% filter(E_Coli > 2000)
water %>% filter(E_Coli > 100)
water %>% filter(Enterococci > 100)

#9 E_Coli Samples are over 100.
#As a whole, it follows the guidelines of 90% of E_Coli count is under 100. Being 90.625%
#However, if we group sample by Beach, 1 - 1/8 is less than 90%, meaning all 6 listed beaches would be breaking guidelines.
#2 Enterococci samples over 100. 96 Samples. Thus, 2/96 Samples = 2.1%.
#100% - 2.1% = 97.9% are within the limit, Guidelines are followed for Enterococci, even if we group by beach.
```


f. Using the full water dataset use facet_wrap to show Enterococci time plots separately for all the beaches. Show the the guideline value.
```{r}
#ggplot(data = water, aes(x = Date, y = Enterococci)) + geom_point() + facet_wrap(vars(~Beach))

water6 <- water %>% pivot_longer(c(Enterococci), names_to = "Enterococci", values_to = "Readings")
ggplot(data = water6, aes(x = Date, y = Readings, color = Enterococci)) + 
  geom_point() + geom_line() + geom_hline(yintercept = 100, color = "magenta")+facet_wrap(~Beach)
```

g. Repeat f. for E_Coli. 
```{r}
#Guideline values: Less than 2000 per 100ml for 95% of all samples.
#Guideline values: Less than 100 per 100ml for 90% of all samples.
water6 <- water %>% pivot_longer(c(E_Coli), names_to = "E_Coli", values_to = "Readings")
ggplot(data = water6, aes(x = Date, y = Readings, color = E_Coli)) + 
  geom_point() + geom_line() + geom_hline(yintercept = 100, color = "magenta")+geom_hline(yintercept = 2000, color = "magenta")+facet_wrap(~Beach)
```

h. (Advanced, optional). Show the Enterococci and E_Coli time plots overlaid for all the beaches. Use colour to distinguish the two y variables.
```{r}
water5 <- water %>% pivot_longer(c(Enterococci, E_Coli), names_to = "Bacteria", values_to = "Readings")
ggplot(data = water5, aes(x = Date, y = Readings, color = Bacteria)) + 
  geom_point() + geom_line() + geom_hline(yintercept = 100, color = "magenta")+facet_wrap(~Beach)
```

i. Write a few sentences about the water quality in the beaches.

Portmarnock had an insanely high E_Coli count on 2013-08-19. Cause should be investigated.
Howth - Claremont and Balbriggan are the dirtiest beaches on the list and should be properly cleaned.
Portrane - Tower Bay had the lowest sample numbers among all the beaches, making it the cleanest of the bunch.
Skerries, Rush - South Beach, Sutton & Malahide had all samples taken to be under guidelines, making them safe beaches.
2013-08-19 is a date of interest. 3 samples from those dates break guidelines and the rise in Bacteria count is noticable even on clean beaches like Portrane - Burrow (The Brook)

```


## Question 4

a. Make histograms of both Enterocci and E_Coli. Do you need the histogram to start at zero? Choose good colour and fills and a binwidth of 10. Arrange your two histograms side by side.
```{r,  out.width="50%", fig.show='hold'}
require(gridExtra)
plot1<- ggplot(water, aes(x= Enterococci)) + geom_histogram(fill="red", color="black", binwidth=10)
plot2 <- ggplot(water, aes(x= E_Coli))      + geom_histogram(fill="lightblue", color="navy", binwidth=10)
grid.arrange(plot1, plot2, ncol=2)

```


b. Make boxplots of both Enteroccoci and E_Coli. Arrange your two boxplots side by side.
```{r,  out.width="50%", fig.show='hold'}
plot3 <- ggplot(water, aes(y= Enterococci)) + geom_boxplot(fill="red", color="black")
plot4 <- ggplot(water, aes(y = E_Coli)) + geom_boxplot(fill="lightblue", color="navy")
grid.arrange(plot3, plot4, ncol=2)

```


c. Now make side by side boxplots of Enterococci and E_Coli. For this you will need to put the data into long form:
```{r}

waterlong <- water %>% pivot_longer(c(Enterococci, E_Coli), names_to="Bacteria", values_to="Reading")
ggplot(waterlong, aes(y = Bacteria)) + geom_boxplot(fill="lightblue", color="navy") + facet_wrap(vars(Date))

```

d.Write a few sentences with your findings. Mention skewness and outliers in your answer.

A boxplot allows us to easily see all the outliers for the Bacteria. Using the boxplot for E_Coli, we see all outliers go past the 100 guideline. This isn't the case for Enterococci. Same amount of Outliers each.
Only one whisker for each boxplot.
E_Coli has the bigger box.


